Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2014-08-26 paper.tex
Added a couple papers by Raskin/Moll/Kasen/etc. and thus concluded the first draft of the DD theory history section.

The first three-dimensional simulations of double degenerate systems were performed by \citet{benz:1990}, who used the smoothed particle hydrodynamics (SPH) method to simulate the merger process. They found that if the lower-mass secondary was close enough to the primary to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope on the primary, with a centrifugally-supported accretion disk surrounding the core and envelope. They noted that the envelope temperature was hot enough to generate carbon fusion but neglected reactions in their simulation. Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009} corroborated this result. This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community. \cite{mochkovitch_livio:1990} and \cite{livio:2000} observed that turbulent viscosity would be sufficiently large for angular momentum to be removed from the disk at a rate high enough to generate these troublesome accretion timescales. Based on all of this evidence, the review of \cite{hillebrandtniemeyer2000} argued that the model was only viable if the accretion-induced collapse problem could be avoided. Later work by \cite{shen:2012} and \cite{schwab:2012} used a more detailed treatment of the viscous transport in the outer regions of the remnant and found that while the centrifugally supported envelope would be converted into hot envelope material on a viscous timescale, their simulations still led to off-center carbon burning. \cite{vankerkwijk:2010} argued that equal-mass mergers would lead to the conditions necessary for carbon detonation in the center of the merged object, but \cite{shen:2012} questioned this for similar reasons related to how viscous transport would convert rotational motion into pressure support. \cite{zhu:2013} followed this with an expanded parameter space study and argued that many of their carbon-oxygen systems had the potential to detonate. The study of the long-term evolution of the remnants is thus still an open subject of research.

@@

The most recent developments give some areas of agreement and some of remaining tension. \cite{pakmor:2012} performed a merger scenario with better treatment of the initial conditions and also found that the merger process happened over more than ten orbits. Nevertheless, they still found that a carbon-oxygen detonation would occur, in line with their earlier results. On the other hand, \cite{dan:2012} and \cite{dan:2014} performed a large sweep of the parameter space for merger pairs and found that pure carbon-oxygen systems would generally not lead to detonations and violent mergers. They did find that for systems containing helium, many of them would detonate and potentially lead to SNe Ia, either through the aforementioned instabilities in the accretion stream, or during the contact phase, similar to the violent carbon-oxygen mergers. \cite{pakmor:2013} added a thin helium shell on their primary white dwarf, and found that this robustly led to a detonation of the white dwarf. Thus the preliminary consensus seems to be that systems containing helium could robustly lead to events resembling SNe Ia, while there is still debate about larger-mass systems containing only carbon and oxygen.

>>>>

The first three-dimensional simulations of double degenerate systems were performed by \citet{benz:1990}, who used the smoothed particle hydrodynamics (SPH) method to simulate the merger process. They found that if the lower-mass secondary was close enough to the primary to begin mass transfer on a dynamical time scale, the secondary completely disrupted and formed a hot envelope on the primary, with a centrifugally-supported accretion disk surrounding the core and envelope. They noted that the envelope temperature was hot enough to generate carbon fusion but neglected reactions in their simulation. Later simulations of similar setups \citep{rasio_shapiro:1995,yoon:2007,loren-aguilar:2009,raskin:2012} corroborated this result. This validation of the prediction of \cite{tutukov_yungelson:1979} did not instill confidence in the community. \cite{mochkovitch_livio:1990} and \cite{livio:2000} observed that turbulent viscosity would be sufficiently large for angular momentum to be removed from the disk at a rate high enough to generate these troublesome accretion timescales. Based on all of this evidence, the review of \cite{hillebrandtniemeyer2000} argued that the model was only viable if the accretion-induced collapse problem could be avoided. Later work by \cite{shen:2012} and \cite{schwab:2012} used a more detailed treatment of the viscous transport in the outer regions of the remnant and found that while the centrifugally supported envelope would be converted into hot envelope material on a viscous timescale, their simulations still led to off-center carbon burning. \cite{vankerkwijk:2010} argued that equal-mass mergers would lead to the conditions necessary for carbon detonation in the center of the merged object, but \cite{shen:2012} questioned this for similar reasons related to how viscous transport would convert rotational motion into pressure support. \cite{zhu:2013} followed this with an expanded parameter space study and argued that many of their carbon-oxygen systems had the potential to detonate. The study of the long-term evolution of the remnants is thus still an open subject of research.

@@

The most recent developments give some areas of agreement and some of remaining uncertainty. \cite{pakmor:2012} performed a merger scenario with a $1.1\ M_\odot$ and $0.9\ M_\odot$ setup, with better treatment of the initial conditions, and also found that the merger process happened over more than ten orbits. Nevertheless, they still found that a carbon-oxygen detonation would occur, in line with their earlier results. \cite{moll:2014} was also able to find a detonation in a similarly massive system. \cite{dan:2012} and \cite{dan:2014} performed a large sweep of the parameter space for merger pairs and found that pure carbon-oxygen systems would generally not lead to detonations and violent mergers except for the most massive systems. They did find that for systems containing helium, many of them would detonate and potentially lead to SNe Ia, either through the aforementioned instabilities in the accretion stream, or during the contact phase, similar to the violent carbon-oxygen mergers. \cite{pakmor:2013} added a thin helium shell on their primary white dwarf, and found that this robustly led to a detonation of the white dwarf. Thus the preliminary consensus seems to be that systems containing helium could robustly lead to events resembling SNe Ia, as well as very massive carbon-oxygen binaries.
